---
title: "cross_validation"
author: "Ravi Brenner"
date: "2024-11-12"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(SemiPar)

set.seed(1)
```

Look at LIDAR data
```{r}
data("lidar")

lidar_df <- lidar |>
  as_tibble() |>
  mutate(id = row_number())
```

```{r}
lidar_df |>
  ggplot(aes(x = range, y = logratio)) + 
  geom_point()
```

## Try to do CV
Compare 3 models - linear, smooth, wiggly

construct training and testing df
```{r}
train_df <- sample_frac(lidar_df, size = 0.8)
test_df <- anti_join(lidar_df, train_df, by = "id")
```

look at these
```{r}
train_df |>
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() +
  geom_point(data = test_df, color = "red")
```

fit three models
```{r}
linear_mod <- lm(logratio ~ range, data = train_df)
smooth_mod <- gam(logratio ~ s(range), data = train_df)
wiggly_mod <- gam(logratio ~ s(range, k = 30), sp = 10e-6, data = train_df)
```

Look at these fits
```{r}
train_df |>
  add_predictions(linear_mod) |>
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df |>
  add_predictions(wiggly_mod) |>
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df |>
  add_predictions(smooth_mod) |>
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  #geom_point(data = test_df, color = "red") + 
  geom_line(aes(y = pred), color = "red")
```

compare numerically using RMSE (want lowest value)
```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

## repeat the train/test split

```{r}
cv_df <- 
  crossv_mc(lidar_df, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

```{r, eval=FALSE}
cv_df |>
  pull(train) |>
  nth(4) 
```

Fit models, extrans RMSEs
```{r}
cv_res_df <- cv_df |>
  mutate(
    linear_mod = map(train, \(x) lm(logratio ~ range, data = x)),
    smooth_mod = map(train, \(x) gam(logratio ~ s(range), data = x)),
    wiggly_mod = map(train, \(x) gam(logratio ~ s(range, k=30), sp = 10e-6, data = x))
    ) |>
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse),
    rmse_smooth = map2_dbl(smooth_mod, test, rmse),
    rmse_wiggly = map2_dbl(wiggly_mod, test, rmse)
  )
```

Look at RMSE distribution

```{r}
cv_res_df |>
  select(starts_with("rmse")) |>
  pivot_longer(everything(),
               names_to = "model",
               values_to = "rmse",
               names_prefix = "rmse_") |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

## Weights of children example
```{r}
child_df <- read_csv("data/nepalese_children.csv") |>
  mutate(weight_ch7 = (weight > 7) * (weight -7))
```

Look at data
```{r}
child_df |>
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = 0.5)
```

Fit some models
```{r}
linear_mod = lm(armc ~ weight, data = child_df)
pwl_mod = lm(armc ~ weight + weight_ch7, data = child_df)
smooth_mod = gam(armc ~ s(weight), data = child_df)
```

Look at models
```{r}
child_df |>
  add_predictions(linear_mod) |>
  ggplot(aes(x = weight, y =  armc)) + 
  geom_point(alpha = 0.5) + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
child_df |>
  add_predictions(pwl_mod) |>
  ggplot(aes(x = weight, y =  armc)) + 
  geom_point(alpha = 0.5) + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
child_df |>
  add_predictions(smooth_mod) |>
  ggplot(aes(x = weight, y =  armc)) + 
  geom_point(alpha = 0.5) + 
  geom_line(aes(y = pred), color = "red")
```

CV to select models
```{r}
cv_df <- crossv_mc(child_df, 100) |>
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
```

Apply models and extract RMSE
```{r}
cv_res_df <- cv_df |>
  mutate(linear_mod = map(train, \(x) lm(armc ~ weight, data = x)),
         pwl_mod = map(train, \(x) lm(armc ~ weight + weight_ch7, data = x)),
         smooth_mod = map(train, \(x) gam(armc ~ s(weight), data = x))) |>
  mutate(rmse_linear = map2_dbl(linear_mod, test, rmse),
         rmse_pwl = map2_dbl(linear_mod, test, rmse),
         rmse_smooth = map2_dbl(smooth_mod, test, rmse))
```

```{r}
cv_res_df |>
  select(starts_with("rmse")) |>
  pivot_longer(everything(),
               names_to = "model",
               values_to = "rmse",
               names_prefix = "rmse_") |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

