---
title: "Linear models"
author: "Ravi Brenner"
date: "2024-11-07"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

Loading nyc airbnb data
```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type) |>
  mutate(borough = fct_infreq(borough),
         room_type = fct_infreq(room_type))
```

## Fit some models

Basic linear model fit
```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

summary(fit)

names(summary(fit))

coef(fit)

fit |>
  broom::tidy() |>
  select(term, estimate, p.value)

fit |>
  broom::glance()
```

Try a bit more complex model
```{r}
fit <- lm(price ~ stars + borough, data = nyc_airbnb)

fit |> 
  broom::tidy() |>
  select(term, estimate, p.value) |>
  mutate(term = str_replace(term,"borough", "Borough: ")) |>
  knitr::kable(digits = 3)
```


## Some diagnostics

(backtrack to some EDA)
```{r}
nyc_airbnb |>
  ggplot(aes(x = stars, y = price)) + 
  geom_point() + 
  stat_smooth(method ="lm")
```


Most diagnostics use residuals

```{r}
modelr::add_residuals(nyc_airbnb, fit) |>
  ggplot(aes(x = borough, y = resid)) + 
  geom_violin() +
  ylim(-200,500)
```

(Some issues here. Highly skewed residuals, nonconstant variance by bourough. May be violating some model assumptions).

Residuals against stars
```{r}
modelr::add_residuals(nyc_airbnb, fit) |>
  ggplot(aes(x = stars, y = resid)) + 
  geom_point() 
```


Residuals against fitted values
```{r}
nyc_airbnb |>
  modelr::add_residuals(fit) |>
  modelr::add_predictions(fit) |>
  ggplot(aes(x = pred, y = resid)) + 
  geom_point()
```

## comparing models
```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```

```{r}
fit_null |>
  broom::tidy()

fit_alt |>
  broom::tidy()

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

## Do effects differ across bouroughs?

First, could use a lot of interactions
```{r}
nyc_airbnb |>
  lm(price ~ stars * borough + room_type * borough, data = _) |>
  broom::tidy()
```

Could just fit separate models
```{r}
nyc_airbnb |>
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()

nyc_airbnb |>
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()
```

Get fancy and use list columns
```{r}
nyc_airbnb |>
  nest(data = -borough) |>
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
    ) |>
  unnest(results) |>
  select(borough, term, estimate) |>
  pivot_wider(names_from = term,
              values_from = estimate)
```

What about room type across manhattan neighborhoods?

```{r}
nyc_airbnb |>
  filter(borough == "Manhattan",
         neighborhood == "Chinatown") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy() 

nyc_airbnb |>
  filter(borough == "Manhattan",
         neighborhood == "Chelsea") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy() 
```

But can do this systematically for all neighborhoods
```{r}
manhattan_neighborhood_fit_df <- nyc_airbnb |>
  filter(borough == "Manhattan") |>
  nest(data = -c(borough, neighborhood)) |>
  mutate(model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
         results = map(model, broom::tidy)) |>
  select(neighborhood, results) |>
  unnest(results) 
```

Look at effect of room type
```{r}
manhattan_neighborhood_fit_df |>
  filter(str_detect(term, "room_type")) |>
  ggplot(aes(x = term, y = estimate)) + 
  geom_boxplot()
```

```{r}
manhattan_neighborhood_fit_df |>
  filter(str_detect(term, "room_type"),
         estimate > 0) 

nyc_airbnb |> filter(neighborhood == "NoHo",
                     room_type == "Shared room")
```


